Exploring vectorization and unrolling with SMT

### 2022-05-25 article/article.tex
atricle with curves and more curves

\begin{figure}
  \begin{lstlisting}
for (int i=0;i<BLOCKI;i+=ui) {                
  for (int j=0;j<BLOCKJ;j+= uj*nv) {           
    { ii in [0,ui],jj in [0,uj]:
        mipp::Reg<TYPE> c_ii,jj;
        c_{ii,jj}.load(&C[(i + ii )*BLOCKJ + j + jj *nv]);
    }                
    for (int k=0;k<BLOCKK;k+= uk ) {            
      { ii in [0,ui],kk in [0,uk]:
          mipp::Reg<TYPE> a{{ ii }}{{ kk }};
          a_{ii,kk} = mipp::set1<TYPE>(A[(i + ii )*BLOCKK+k+ kk ]);
      }
      { jj in [0,uj],kk in [0,uk]:
          mipp::Reg<TYPE> b{{ kk }}{{ jj }};
          b_{kk,jj}.load(&B[(k+ kk )*BLOCKJ + j + jj*nv]);
      }    
      { ii in [0,ui], jj in [0,uj] :
          c_{ii,jj} = { k in [0,uk]: mipp::fmadd(a_ii,kk, b_kk,jj, } , c_ii,jj  { k in [0,uk]: ) };
      }
    }
    { ii in [0,ui],jj in [0,uj]:
        c_ii,jj.store(&C[(i+ ii )*BLOCKJ + j + ( jj )*nv]);      
    }
  }
}
  \end{lstlisting}
  \caption{Matrix product vectorized and unrolled\label{fig:gemm}. \texttt{in} constructs correspond to unrolled statements/expressions. }
\end{figure}

@@

\textbf{For blocked  Gemm, it shows that SMT brings $8\%$ of performance, and the codes with good performance for SMT have good performance on 1 thread.} 
\section{Related Works}
\section{Conclusion}
\label{sec:conclusion}

>>>>

\begin{figure}
  \begin{lstlisting}
for (int i=0;i<BLOCKI;i+=ui) {                
  for (int j=0;j<BLOCKJ;j+= uj*nv) {           
    { ii in [0,ui],jj in [0,uj]:
        mipp::Reg<TYPE> c_ii,jj;
        c_{ii,jj}.load(&C[(i + ii )*BLOCKJ + j + jj *nv]);
    }                
    for (int k=0;k<BLOCKK;k+= uk ) {            
      { ii in [0,ui],kk in [0,uk]:
          mipp::Reg<TYPE> a{ i }{ k };
          a_{ii,kk} = mipp::set1<TYPE>(A[(i + ii )*BLOCKK+k+ kk ]);
      }
      { jj in [0,uj],kk in [0,uk]:
          mipp::Reg<TYPE> b{ k }{{ jj }};
          b_{kk,jj}.load(&B[(k+ kk )*BLOCKJ + j + jj*nv]);
      }    
      { ii in [0,ui], jj in [0,uj] :
          c_{ii,jj} = { k in [0,uk]: mipp::fmadd(a_ii,kk, b_kk,jj, } , c_ii,jj  { k in [0,uk]: ) };
      }
    }
    { ii in [0,ui],jj in [0,uj]:
        c_ii,jj.store(&C[(i+ ii )*BLOCKJ + j + ( jj )*nv]);      
    }
  }
}
  \end{lstlisting}
  \caption{Matrix product vectorized and unrolled\label{fig:gemm}. \texttt{in} constructs correspond to unrolled statements/expressions. }
\end{figure}

@@

\textbf{For blocked  Gemm, it shows that SMT brings $8\%$ of performance, and the codes with good performance for SMT have good performance on 1 thread.}

%=========== Ajouté par Edgar (donc j'ai peut être fait des erreurs ou fautes de frappes) ================

\subsection{Matrix-vector operations performances on AMD RYZEN 5}

The AMD RYZEN 5 processor has 16 vector registers and 256 bits in a vector, so it can store 4 floats.\\

In the following demonstrations, $i$ and $j$ are the unrolling degree on one dimensions of the matrix.

\begin{figure*}[h!]
  \includegraphics[width=\textwidth]{../autobench/datas/axya/axya_ui_1_32_uk_1_32.pdf}
  \caption{Performance of the sum of product $A\times x+y\times A$, with $A$ a $1024\times1024$ matrix and $x,y$ two vectors of length $1024$. Type of data is float.}
\end{figure*}

Now the Figure \ref{fig:matvec_comp} shows the performances differences of a matrix vector product with different versions of the product algorithm. The two algorithms are described on Figure \ref{fig:matvec_algo}.

\begin{figure}
\begin{tabular}{ccc}
\begin{lstlisting}[linewidth=0.45\textwidth]
for (int k=0;k<BLOCKK;k+=uk) {
  { kk in [0,uk]:
    mipp::Reg<TYPE> b0_{ kk };
  }
  { kk in [0,uk]:
    b0_{ kk }=mipp::set1<TYPE>(B[k+ {{ kk } ]);
  }
  for (int i=0;i<BLOCKI;i+=ui*nv) {         
    { ii in [0,ui]
      mipp::Reg<TYPE> c{ ii };
    }
    { kk in [0,uk]:
      { ii in [0,ui]:
        mipp::Reg<TYPE> a{ ii }_{ kk };
      }
    }
    { ii in [0,ui]:
      c{ ii }.load(&C[i+{ ii }*nv]);
    }
    { kk in [0,uk]:
      { ii in [0,ui]:
        a{ ii }_{ kk }.load(&A[BLOCKI*(k+ { kk } ) + i+{ ii }*nv]);
      }
    }
    { kk in [0,uk]:
      { ii in [0,ui]:
        c{ ii } = mipp::fmadd(a{ ii }_{ kk }, b0_{ kk },c{ ii });
      }
    }
    { ii in [0,ui]:
      c{ ii }.store(&C[i+{ ii }*nv]);
    }
  }
}

\end{lstlisting}
&
\hspace{0.12\textwidth}
&
\begin{lstlisting}[linewidth=0.45\textwidth]
for (int i=0;i<BLOCKI;i+=ui*nv) {
    { ii in [0,ui]
      mipp::Reg<TYPE> c{ i }_0;
    }
    { ii in [0,ui]
      c{ i }_0.load(&C[i+ { i }*nv ]);
    }
    for (int k=0;k<BLOCKK;k+=uk) {         
      { ii in [0,ui]
        { kk in [0,uk]:
          mipp::Reg<TYPE> a{ i }_{ k };
        }
      }
      { kk in [0,uk]:
        mipp::Reg<TYPE> b{ k };
      }
      { kk in [0,uk]:
        b{ k }=mipp::set1<TYPE>(B[k+{ k }]);
      }
      { ii in [0,ui]
        { kk in [0,uk]:
          a{ i }_{ k }.load(&A[BLOCKI*(k+ { k } ) + i+{ i }*nv]);
        }
      }
      { ii in [0,ui]
        { kk in [0,uk]:
          c{ i }_0 = mipp::fmadd(a{ i }_{ k }, b{ k },c{ i }_0);
        }
      }
    }
    { ii in [0,ui]
      c{ i }_0.store(&C[i+{ i }*nv]);
    }
  }   
  \end{lstlisting}
  \\
\end{tabular}

  \caption{Two different matrix-vector product algorithm vectorized and unrolled.}
  \label{fig:matvec_algo}
\end{figure}

\begin{figure*}[h!]
  \begin{subfigure}[h]{0.5\textwidth}  
  \includegraphics[width=\textwidth]{../autobench/datas/matvec/matvec_ui_1_16_uk_1_16.pdf}
  \caption{Store every loop step}
  \end{subfigure}
  \begin{subfigure}[h]{0.5\textwidth}  
  \includegraphics[width=\textwidth]{../autobench/datas/matvec_fewstores/matvec_fewstores_ui_1_32_uk_1_32.pdf}
  \caption{Broadcast every loop step}
  \end{subfigure}
  \begin{subfigure}[h]{0.5\textwidth}  
  \includegraphics[width=\textwidth]{../autobench/datas/matvec/threading/matvec_ui_1_16_uk_1_16_hyper_threading.pdf}
  \caption{Store every loop step (hyper threading)}
  \end{subfigure}
  \begin{subfigure}[h]{0.5\textwidth}  
  \includegraphics[width=\textwidth]{../autobench/datas/matvec_fewstores/threading/matvec_fewstores_ui_1_16_uk_1_16_hyper_threading.pdf}
  \caption{Broadcast every loop step (hyper threading)}
  \end{subfigure}
  \begin{subfigure}[h]{0.5\textwidth}  
  \includegraphics[width=\textwidth]{../autobench/datas/matvec/threading/matvec_ui_1_16_uk_1_16_multi_threading.pdf}
  \caption{Store every loop step (multi threading)}
  \end{subfigure}
  \begin{subfigure}[h]{0.5\textwidth}  
  \includegraphics[width=\textwidth]{../autobench/datas/matvec_fewstores/threading/matvec_fewstores_ui_1_16_uk_1_16_multi_threading.pdf}
  \caption{Broadcast every loop step (multi threading)}
  \end{subfigure}

  \caption{Comparison of $A\times x$ performance, with $A$ a $1024\times1024$ matrix and $x$ a vector of length $1024$. Type of data is float.}
  \label{fig:matvec_comp}
\end{figure*}

A lot of points on those graphs are red because every float of the matrix in loaded only one time, so in the assembly code, the \texttt{fmadd} instruction is directly loading the vector register, like this:

\begin{center}
  \texttt{vfmadd231ps -0x40(\%rdi),\%ymm9,\%ymm2}\\
  \texttt{vfmadd231ps -0x20(\%rdi),\%ymm9,\%ymm1}
\end{center}

Thus, only $i+j$ vector register are physically used, while there is theoretically $i+j+i*j$ vector registers used at the same time. This behavior allows a greater unrolling without using the spill.

%===============================

\section{Related Works}
\section{Conclusion}
\label{sec:conclusion}
